---
title: "P805_hw1_lx2346"
author: "linshan"
date: "2024-09-12"
output: github_document
---

# Problem 1
## short description of the penguins dataset 
Load the penguins dataset 
```{r}
data("penguins", package = "palmerpenguins")
```

**The penguins data set has `r nrow(penguins)` rows and `r ncol(penguins)` columns.**\
\
**The variables in the dataset are `r names(penguins)`.**\
\
The key varibles and their values in the dataset include:\
* sex\
* year\
* species : `r unique(penguins$species)`\
* island : `r unique(penguins$island)`\
* bill_length_mm : length of penguins' bill, ranging from `r min(penguins$bill_length_mm,na.rm = TRUE)` to `r max(penguins$bill_length_mm, na.rm = TRUE)` mm.\
* bill_depth_mm : length of penguins' bill, ranging from `r min(penguins$bill_depth_mm,na.rm = TRUE)` to `r max(penguins$bill_depth_mm, na.rm = TRUE)` mm.\
* flipper_length_mm : length of penguins' flipper, ranging from `r min(penguins$flipper_length_mm,na.rm = TRUE)` to `r max(penguins$flipper_length_mm, na.rm = TRUE)` mm.\
* body_mass_g : ranging from `r min(penguins$body_mass_g,na.rm = TRUE)` to `r max(penguins$body_mass_g, na.rm = TRUE)` g.\
\
**The mean of the flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.**\
\
## Scatterplot of flipper_length_mm vs bill_length_mm
```{r}
library(ggplot2)
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm))+
  geom_point(aes(color = species))
ggsave("scatter_fli_vs_bill.png")
```

# Problem 2
## create a data frame
```{r}
library(tidyverse)
char_ori <- c("a","b","c")
fac_ori <- factor(c("low","medium","high"))
data_framee <- tibble(
  random_var = rnorm(10),
  logi_vec = random_var > 0,
  char_vec = sample(char_ori,size = 10,replace = TRUE),
  fac_vec = sample(fac_ori, size = 10, replace = TRUE))
```
From above, a data frame with 4 columns are constructed.

## calculate the mean of each varibale
mean of the random sample: `r mean(pull(data_framee,random_var))`.\
mean of the logical vector: `r mean(pull(data_framee,logi_vec))`.\
mean of the character vector: `r mean(pull(data_framee,char_vec))`.\
mean of the factor vector:  `r mean(pull(data_framee,fac_vec))`.\

It seems that only the random varible and the logical vector have the means. The rest of the two vectors' mean function doesn't work.

### If I use as.numeric function, what will happen next?
```{r results='hide'}
logi_asnum <- as.numeric(pull(data_framee,logi_vec))
char_asnum <- as.numeric(pull(data_framee,char_vec))
fac_asnum <- as.numeric(pull(data_framee,fac_vec))
```
## Calculate the means again.
mean of the logical vector: `r mean(logi_asnum)`.\
mean of the character vector: `r mean(char_asnum)`.\
mean of the factor vector: `r mean(fac_asnum)`.\
**As it is shown in the result, when converted to numeric data type, the mean of the logical vector and factor vector can be calculated.**\
* When converting factors to numeric, as.numeric() first converts the factor levels to integers, such as 1,2 and 3 in my example.\
* When converting logical variables, TRUE will be converted to 1 and FULSE will be converted to 0.\
* However, when the object need to be converted is character, things may get a little complicated.\
- If the character vector represents numbers (eg."1","3"), as.numberic() will convert these characters into numeric values.\
- However, if the character vector does not represent numbers, the function will return NA, then the mean cannot be calculated.




